---
title: "Survey Assessing Risks from AI (SARA) USA 2026"
subtitle: "Pre-registered analysis report"
author: "Generated by analysis/report.Rmd (single reproducible pipeline)"
date: "`r format(Sys.time(), '%d %B %Y, %H:%M %Z')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: cosmo
    css: [fonts_figtree.css, report_theme.css]
params:
  data_dir: "simdata"
  blocks_csv: "../survey/sara/dce_blocks.csv"
  instrument: "../survey/sara_usa.md"
  n_pstar_draws: 4000
  n_boot: 1000
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 4.5)
suppressMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(knitr)
  library(lme4); library(boot); library(MASS, exclude = "select")
})
HERE <- normalizePath(".")
source("report_utils.R")
source("../dce_model_utils.R")   # registered DCE coding + estimator (shared)

DATA   <- params$data_dir
WIDE   <- file.path(DATA, "all_apps_wide.csv")
ORDERS <- file.path(DATA, "orders.csv")
PTIMES <- file.path(DATA, "PageTimes.csv")
TRUTH  <- file.path(DATA, "truth.json")
CACHE  <- file.path(DATA, "cache"); dir.create(CACHE, showWarnings = FALSE)

SIMULATED <- file.exists(TRUTH)
truth <- if (SIMULATED) jsonlite::fromJSON(TRUTH) else NULL

# memoise expensive fits on the input data's fingerprint
memo <- function(name, expr) {
  key <- paste0(name, "_", substr(unname(tools::md5sum(WIDE)), 1, 10), ".rds")
  f <- file.path(CACHE, key)
  if (file.exists(f)) return(readRDS(f))
  v <- force(expr); saveRDS(v, f); v
}

spec  <- load_instrument(params$instrument)
items <- spec_items(spec)
lab   <- function(id) item_labels(spec, items, id)
```

`r if (SIMULATED) '> **SIMULATED DATA.** This render uses data from analysis/simulate_data.py (format-identical to the oTree server export; see analysis/check_export_schema.py). Every number below is synthetic. The same document, pointed at the real export, is the registered analysis. A parameter-recovery appendix closes the report.'`

This report implements **every registered analysis in `PREREGISTRATION.md`**
(§3 exclusions and flags, §4 estimands, §5 design provenance, §6 MRP, §7
benchmarks, §8 negative commitments). Item text, scale labels, and option
codings are read from the instrument (`survey/sara_usa.md`) — the single
source of truth — never restated here.

```{r prep}
# ── Shared data preparation (used by every section below) ────────────
d_all  <- read_wide(WIDE)
blocks <- read.csv(params$blocks_csv, check.names = FALSE)
d_all  <- add_flags(d_all, blocks)

idx <- as.integer(d_all[["participant._index_in_pages"]])
maxidx <- as.integer(d_all[["participant._max_page_index"]])[1]
d_all$declined <- num(d_all$consent, keep_optout = TRUE) == 2
att_req <- which(lab("m3_att_bioweapons") == "Much less strict")
a1 <- num(d_all$m3_att_bioweapons, keep_optout = TRUE)
a2 <- num(d_all$m3_att_nuclear,   keep_optout = TRUE)
d_all$screened <- !is.na(a1) & !is.na(a2) & a1 != att_req & a2 != att_req
d_all$completed <- idx > maxidx
d_all$partial <- !d_all$declined & !d_all$screened & !d_all$completed
d_all$wave <- d_all$session.label

# Analysis sample: consented, not screened out (registered exclusion §3.1),
# and completed the survey (incomplete Prolific submissions are returned and
# replaced; their partial data are not analysed).
d <- d_all %>% filter(!declined, !screened, completed)

# §3.6 speed flag: briefing dwell below the pilot (wave 1) 5th percentile.
bd <- briefing_dwell(PTIMES)
d <- d %>% left_join(bd, by = c("participant.code" = "participant_code"))
pilot_floor <- quantile(d$brief_dwell[d$wave == "wave1"], 0.05, na.rm = TRUE)
d$flag_speed <- !is.na(d$brief_dwell) & d$brief_dwell < pilot_floor

# Disclosure arm, numeracy quartile, ladder matrix, scope measures
d$info <- num(d$info_arm, keep_optout = TRUE) == 1
d$bnt_q <- bnt_quartile(d, items)
ladder_ids <- c("m4c_single", "m4c_100", "m4c_1m", "m4c_800m", "m4c_extinction")
ladder_lab <- lab("m4c_single")
ladder_short <- c(m4c_single = "1 death", m4c_100 = "100 deaths",
                  m4c_1m = "1M deaths", m4c_800m = "10% of humanity",
                  m4c_extinction = "Extinction")   # axis labels only
rungs <- sapply(ladder_ids, function(i) num(d[[i]]))
d$scope_slope <- rungs[, "m4c_extinction"] - rungs[, "m4c_single"]
d$scope_flat  <- apply(rungs, 1, function(r) all(!is.na(r)) && length(unique(r)) == 1)

# Superintelligence module DVs. Continuous score = pro-ban minus the
# reverse-keyed anti-ban sentiment: pro = 5 - support (higher = support the
# ban; "Don't know" -> missing); anti-ban sentiment = 6 - anti (agree5).
d$cell <- num(d$muskan_cell, keep_optout = TRUE)
mk_sup <- num(d$muskan_support); mk_sup[mk_sup == 5] <- NA
mk_anti <- num(d$muskan_support_anti)
d$mk_score <- (5 - mk_sup) - (6 - mk_anti)
d$mk_top2 <- num(d$muskan_support) %in% 1:2
d$mk_top2[is.na(num(d$muskan_support))] <- NA
d$anti_top2 <- mk_anti %in% 1:2

# DCE machinery shared by §4, §8, and the headline summary
resp <- d %>% transmute(code = participant.code,
                        dce_block = num(dce_block, keep_optout = TRUE),
                        across(all_of(paste0("dce_", 1:10)),
                               ~ num(., keep_optout = TRUE)))
fit_dce <- function(rows, tag) memo(paste0("dce_", tag), {
  long <- dce_build_long(as.data.frame(rows), params$blocks_csv)
  set.seed(42)                       # registered fixed seed for the multistart
  dce_estimate(long, numMultiStarts = 5)
})
pstar_draws <- function(m, ndraws) {
  keep <- MAIN_PARS
  dr <- MASS::mvrnorm(ndraws, coef(m)[keep], vcov(m)[keep, keep])
  grid <- expand.grid(severity = names(SEV_TO_LOGN),
                      benefit = c("Modest", "Major", "Transformative"),
                      competition = c("Other countries are ahead",
                                      "The US keeps pace", "The US is ahead"),
                      stringsAsFactors = FALSE)
  res <- lapply(seq_len(nrow(grid)), function(i) {
    g <- grid[i, ]
    v_opt <- (dr[, "logn"] * SEV_TO_LOGN[[g$severity]]
              + dr[, "ext"] * (g$severity == EXT_LEVEL)
              + dr[, "ben_major"] * (g$benefit == "Major")
              + dr[, "ben_transf"] * (g$benefit == "Transformative")
              + dr[, "comp_pace"] * (g$competition == "The US keeps pace")
              + dr[, "comp_lead"] * (g$competition == "The US is ahead"))
    logp_star <- (dr[, "no_choice"] - v_opt) / dr[, "logp"]
    p <- 10^logp_star
    cbind(g, t(quantile(p, c(0.5, 0.025, 0.975))))
  })
  out <- bind_rows(res)
  names(out)[4:6] <- c("median", "lo", "hi")
  out
}
```

# Headline results at a glance

Every value below is computed by the registered code in its own section
(linked); this table is presentation only.

```{r headlines}
m_dce <- fit_dce(resp, "primary")
ps_ext <- pstar_draws(m_dce, params$n_pstar_draws) %>%
  filter(severity == EXT_LEVEL, benefit == "Major",
         competition == "The US keeps pace")
c5 <- d %>% filter(cell == 5)
c5ci <- prop_ci(sum(c5$mk_top2, na.rm = TRUE), sum(!is.na(c5$mk_top2)))
ext_r <- rungs[, "m4c_extinction"]
headlines <- tribble(
  ~headline, ~value, ~section,
  "Analysis sample", sprintf("%d completers (of %d rows exported)", nrow(d), nrow(d_all)), "§1",
  "Stated highest acceptable extinction chance (median)",
    sprintf("%s; %s answer 'Never allowed'", ladder_lab[ordinal_median(ext_r)],
            pct(mean(ext_r == 9, na.rm = TRUE))), "§3",
  "Scope-insensitive (flat ladder)", pct(mean(d$scope_flat, na.rm = TRUE)), "§3",
  "DCE p* at extinction (Major benefit, US keeps pace)",
    sprintf("%s per year [95%%: %s to %s]", fmt_1inN(ps_ext$median),
            fmt_1inN(ps_ext$hi), fmt_1inN(ps_ext$lo)), "§4",
  "Superintelligence ban: considered support (cell 5, top-2-box)",
    sprintf("%s [%s, %s]", pct(c5ci[1]), pct(c5ci[2]), pct(c5ci[3])), "§9",
  "Method 1 vs Method 2 convergence gate", "see §8 (computed on disclosed-arm data)", "§8",
  "Panel-selection benchmark gaps vs Pew", "see §2 - read before any population claim", "§2")
kable(headlines, caption = "Headlines. Point values must be read with the intervals, flags, and caveats in their sections.")
```

# 0. Provenance

```{r provenance}
git <- tryCatch(system("git rev-parse --short HEAD", intern = TRUE), error = function(e) "n/a")
prov <- data.frame(
  input = c("git commit", "all_apps_wide.csv", "dce_blocks.csv", "instrument"),
  value = c(git,
            unname(tools::md5sum(WIDE)),
            unname(tools::md5sum(params$blocks_csv)),
            unname(tools::md5sum(params$instrument))))
kable(prov, caption = "Inputs (md5). The DCE design fielded at every wave is the committed dce_blocks.csv (PREREGISTRATION.md §5: archival design reproducibility).")
```

# 1. Sample flow and registered exclusions (§3)

```{r flow}
flow <- tibble(
  stage = c("Started (rows in export)", "Declined consent",
            "Screened out (both attention checks failed; excluded, §3.1)",
            "Incomplete (returned/timed out; not analysed)",
            "Analysis sample"),
  n = c(nrow(d_all), sum(d_all$declined), sum(d_all$screened),
        sum(d_all$partial), nrow(d)))
kable(flow, caption = "Sample flow. Screened-out participants are replaced on Prolific; their partial data are never analysed.")
kable(count(d, wave), caption = "Analysis sample by fielding wave (oTree session).")
```

```{r flags}
flags <- tribble(
  ~flag, ~rule, ~handling, ~n, ~share,
  "DCE dominated option (§3.2)", "chose the dominated side of task 9 (opt-out is NOT a failure)",
    "retained; sensitivity re-estimate", sum(d$flag_dominated, na.rm = TRUE), mean(d$flag_dominated, na.rm = TRUE),
  "DCE stability (§3.3)", "task 10 choice differs from its exact repeat, task 2",
    "retained; sensitivity re-estimate", sum(d$flag_unstable, na.rm = TRUE), mean(d$flag_unstable, na.rm = TRUE),
  "Sanity anchor (§3.4)", "AI should be less strictly regulated than Everest climbing",
    "descriptive only", sum(d$flag_everest, na.rm = TRUE), mean(d$flag_everest, na.rm = TRUE),
  "Acquiescence (§3.5)", "top-2-box on BOTH the pro-ban and anti-ban items",
    "module sensitivity only", sum(d$flag_acquiescer, na.rm = TRUE), mean(d$flag_acquiescer, na.rm = TRUE),
  "Speed (§3.6)", sprintf("briefing dwell < pilot 5th percentile (= %.0f s)", pilot_floor),
    "sensitivity only", sum(d$flag_speed, na.rm = TRUE), mean(d$flag_speed, na.rm = TRUE))
flags$share <- pct(flags$share)
kable(flags, caption = "Registered data-quality flags (§3). None is an exclusion; scope-insensitivity is a measured outcome, never grounds for dropping (§3.7). No other cleaning was performed.")
```

# 2. Panel-selection benchmarks (§7) — read this before any headline

Two verbatim Pew ATP items were fielded before any treatment. Every
population-level headline in this report must be read next to these gaps.

```{r benchmarks}
pew <- list(
  bench_pew_cncexc = c(0.10, 0.50, 0.38),   # ATP W173, Jun 2025
  bench_pew_aireg  = c(0.21, 0.58, 0.21))   # ATP, Aug 2024
bench <- bind_rows(lapply(names(pew), function(id) {
  x <- num(d[[id]])
  tibble(item = id, option = lab(id),
         sample = as.numeric(table(factor(x, 1:3)) / sum(!is.na(x))),
         pew = pew[[id]])
}))
bench$gap_pp <- 100 * (bench$sample - bench$pew)
kable(bench %>% mutate(sample = pct(sample), pew = pct(pew),
                       gap_pp = sprintf("%+.1f", gap_pp)),
      caption = "Sample vs Pew probability-sample toplines (options in canonical order; display order was rotated per respondent).")
max_gap <- max(abs(bench$gap_pp))
caveat <- max_gap > 10
```

**Largest gap: `r sprintf('%.1f', max_gap)` percentage points.**
`r if (caveat) "**This exceeds the registered 10-point threshold: every population claim in this report carries an explicit calibration caveat, and a raking-to-benchmark sensitivity estimate is reported below.**" else "This is within the registered 10-point threshold, so no calibration caveat is triggered; the gaps are still reported next to every headline by reference to this section."`

```{r raking, eval=caveat}
w_rake <- rake_two_margins(num(d$bench_pew_cncexc), num(d$bench_pew_aireg),
                           setNames(pew$bench_pew_cncexc, 1:3),
                           setNames(pew$bench_pew_aireg, 1:3))
ext <- num(d$m4c_extinction)
amod <- num(d$m2_experts_amodei)
rak <- tibble(
  headline = c("Extinction ladder: % answering 1 in 1,000,000 or stricter (incl. Never)",
               "% Intolerable at Amodei's stated range"),
  raw   = c(mean(ext >= 6, na.rm = TRUE), mean(amod == 2, na.rm = TRUE)),
  raked = c(weighted_share(ext, w_rake, 6:9), weighted_share(amod, w_rake, 2)))
kable(rak %>% mutate(raw = pct(raw), raked = pct(raked)),
      caption = "Registered raking-to-benchmark sensitivity (direction-of-bias check; raked to both Pew items by IPF).")
```

# 3. Method 1 — stated F–N curve (§4.1)

```{r ladder-dist, fig.height=5}
L <- d %>%
  transmute(code = participant.code, info_arm = num(info_arm, keep_optout = TRUE),
            across(all_of(ladder_ids), num)) %>%
  pivot_longer(all_of(ladder_ids), names_to = "severity", values_to = "rung") %>%
  mutate(severity = factor(severity, ladder_ids))
ggplot(filter(L, !is.na(rung)),
       aes(severity, fill = factor(rung, 9:1, rev(ladder_lab)))) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = rev(sara_seq(9))) +   # darkest = strictest
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(labels = ladder_short) +
  labs(y = "share of respondents", x = NULL, fill = "highest acceptable\nannual chance",
       title = "Stated highest acceptable annual chance, by severity") +
  theme_sara()

m1_tab <- L %>% group_by(severity) %>%
  summarise(n = sum(!is.na(rung)),
            median_rung = ordinal_median(rung),
            pct_never = pct(mean(rung == 9, na.rm = TRUE)))
m1_tab$median_label <- ladder_lab[m1_tab$median_rung]
kable(m1_tab, caption = "Per-severity distributions. 'Never allowed' is its own category (reported, never coerced to a number); ordinal medians treat it as the strictest rung.")
```

### Scope sensitivity × numeracy (registered interaction)

```{r scope}
scope_tab <- d %>% filter(!is.na(bnt_q)) %>% group_by(bnt_q) %>%
  summarise(n = n(), pct_flat = pct(mean(scope_flat, na.rm = TRUE)),
            mean_slope = round(mean(scope_slope, na.rm = TRUE), 2))
kable(scope_tab, caption = "Scope sensitivity by Berlin Numeracy quartile (quartile derived from the adaptive path; slope = ordinal position change, extinction minus single death, Never = rung 9).")
cat(sprintf("Overall: %s scope-insensitive (flat across all five severities).\n",
            pct(mean(d$scope_flat, na.rm = TRUE))))
m_scope <- lm(scope_slope ~ bnt_q, data = d)
kable(round(summary(m_scope)$coefficients, 3),
      caption = "Registered interaction test: ladder slope on numeracy quartile (linear).")
```

### Extinction discontinuity (registered contrast)

```{r extinction-contrast}
r800 <- rungs[, "m4c_800m"]; rext <- rungs[, "m4c_extinction"]
ok <- !is.na(r800) & !is.na(rext)
ext_tab <- tibble(
  comparison = c("Stricter at extinction than at 10% of humanity",
                 "Tied", "Laxer at extinction"),
  share = c(mean(rext[ok] > r800[ok]), mean(rext[ok] == r800[ok]),
            mean(rext[ok] < r800[ok])))
kable(ext_tab %>% mutate(share = pct(share)),
      caption = sprintf("Within-person contrast m4c_extinction vs m4c_800m (n = %d; 'Never allowed' handled ordinally as the strictest rung). Test of Schubert, Caviola & Faber (2019): is extinction judged uniquely worse than near-extinction?", sum(ok)))
```

### First-rung sensitivity (anchoring/demand check)

The ladder is one item per page in per-participant random order, so each
respondent's *first-seen* rung is an unanchored between-subjects estimate.
Orders are reconstructed deterministically from `participant.code`
(`derive_orders.py`) — no server-side logging needed.

```{r first-rung}
orders <- read.csv(ORDERS, check.names = FALSE)
d <- d %>% left_join(orders %>% dplyr::select(`participant.code`, severity_ladder_first),
                     by = "participant.code")
fr <- bind_rows(lapply(ladder_ids, function(id) {
  first <- num(d[[id]])[d$severity_ladder_first == id]
  all_  <- num(d[[id]])
  tibble(severity = id,
         n_first = sum(!is.na(first)),
         median_first = ordinal_median(first),
         median_within = ordinal_median(all_),
         p_wilcoxon = tryCatch(
           wilcox.test(first, all_[d$severity_ladder_first != id])$p.value,
           error = function(e) NA))
}))
kable(fr %>% mutate(p_wilcoxon = signif(p_wilcoxon, 2)),
      caption = "First-seen (between-subjects, unanchored) vs full within-person answers, by severity. Registered rule: divergence beyond sampling error is reported, and the first-response estimates are published alongside the within-person headline.")
```

### Willingness to pay (`m5_wtp`) with protest-zero handling

```{r wtp}
wtp <- num(d$m5_wtp); wlab <- lab("m5_wtp")
dk <- which(wlab == "Don't know")
wtp_dist <- tibble(band = wlab, n = as.integer(table(factor(wtp, seq_along(wlab)))))
wtp_dist$share <- pct(wtp_dist$n / sum(!is.na(wtp)))
kable(wtp_dist, caption = "WTP per year to cut catastrophe risk 1-in-20 to 1-in-100 (log-scale bands). 'Don't know' reported separately, excluded from summaries.")

zr <- num(d$m5_wtp_zero_reason); zlab <- lab("m5_wtp_zero_reason")
protest_codes <- grep("should pay|would really cut", zlab)  # registered protest zeros
kable(tibble(reason = zlab, n = as.integer(table(factor(zr, seq_along(zlab))))),
      caption = "Protest-zero probe among $0 respondents. 'Companies/government should pay' and 'money would be wasted' are protest zeros; 'not worth money' and 'cannot afford' are genuine.")

wtp_num <- ifelse(wtp %in% dk, NA, wtp)
protest <- !is.na(zr) & zr %in% protest_codes
sens <- tibble(
  estimate = c("Primary (all zeros retained)",
               "Sensitivity (protest zeros excluded)"),
  n = c(sum(!is.na(wtp_num)), sum(!is.na(wtp_num) & !protest)),
  median_band = wlab[c(ordinal_median(wtp_num),
                       ordinal_median(wtp_num[!protest]))])
kable(sens, caption = "Registered protest-zero sensitivity ('another reason' zeros are retained and reported).")
```

# 4. Method 2 — discrete choice experiment (§4.2, §5)

The registered model: mixed logit (`logitr`), log annual probability with a
random normal coefficient, log severity, an **extinction indicator** (the
discontinuity test), benefit and competition categorical, opt-out constant;
seeded multistart; estimation on tasks 1–8 only (9–10 are quality checks).
Coding, estimator, and long-format construction are shared with the design
pipeline via `dce_model_utils.R`.

```{r dce-fit}
m_dce <- fit_dce(resp, "primary")    # machinery defined in the prep chunk
co <- coef(m_dce); se <- sqrt(diag(vcov(m_dce)))
coef_tab <- tibble(parameter = names(co), estimate = round(co, 3),
                   se = round(se, 3))
if (SIMULATED) coef_tab$truth <- unlist(truth$truth$dce)[
  sub("^sd_logp$", "sd_logp", names(co))]
kable(coef_tab, caption = sprintf(
  "Registered mixed logit, N = %d respondents, estimation tasks only. The 'ext' row is the registered extinction-discontinuity test (disutility beyond log-deaths).%s",
  n_distinct(resp$code), if (SIMULATED) " Truth column: simulation ground truth." else ""))
```

### The estimand p\*(severity, benefit, competition)

p\* is the annual catastrophe probability at which an AI-future option ties
the "keep today's status quo" opt-out — **reported as a distribution with
intervals, never a point** (§8).

```{r pstar, fig.height=5}
ps <- pstar_draws(m_dce, params$n_pstar_draws)
ps_plot <- ps %>%
  filter(competition == "The US keeps pace", median < 1) %>%
  mutate(severity = factor(severity, names(SEV_TO_LOGN)))
ggplot(ps_plot, aes(x = median, y = severity, colour = benefit)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  geom_errorbarh(aes(xmin = pmin(lo, 1), xmax = pmin(hi, 1)), height = 0.25,
                 position = position_dodge(width = 0.5)) +
  scale_colour_manual(values = c(Modest = "#9E9E9E", Major = SARA_RED,
                                 Transformative = "#2E0A12")) +
  scale_x_log10(labels = function(p) fmt_1inN(p)) +
  labs(title = "p*: annual probability at which an AI future ties the status quo",
       subtitle = "medians with 95% intervals; competition held at 'The US keeps pace'; tiers with p* >= 1 (non-binding) omitted",
       x = "acceptable annual probability (log scale)", y = NULL, colour = "benefit") +
  theme_sara()
ps_show <- ps %>%
  mutate(`p* median` = fmt_1inN(median),
         `95% interval` = paste(fmt_1inN(hi), "to", fmt_1inN(lo))) %>%
  dplyr::select(severity, benefit, competition, `p* median`, `95% interval`)
kable(ps_show %>% filter(competition == "The US keeps pace"),
      caption = "p* by severity and benefit (competition: 'The US keeps pace'). 'any chance' = the model puts indifference above p = 1 for that scenario (the tier does not bind). Full grid in the fold below.")
```

<details><summary>Full p\* grid (all 45 scenarios)</summary>
```{r pstar-full}
kable(ps_show)
```
</details>

### Heterogeneity: p\* across the population

The risk slope is a **random** coefficient (registered specification), so p\*
varies across people. The table above is the population-mean respondent;
this view mixes over the estimated slope distribution — how much of the
public sits stricter or laxer than that mean.

```{r pstar-het}
set.seed(42)
het_scenarios <- expand.grid(severity = names(SEV_TO_LOGN)[3:5],
                             benefit = "Major",
                             competition = "The US keeps pace",
                             stringsAsFactors = FALSE)
b <- coef(m_dce)
logp_i <- rnorm(params$n_pstar_draws, b[["logp"]], abs(b[["sd_logp"]]))
logp_i <- logp_i[logp_i < 0]     # a positive slope implies no finite p*; report the share
share_nonneg <- 1 - length(logp_i) / params$n_pstar_draws
het <- bind_rows(lapply(seq_len(nrow(het_scenarios)), function(i) {
  g <- het_scenarios[i, ]
  v_opt <- (b[["logn"]] * SEV_TO_LOGN[[g$severity]]
            + b[["ext"]] * (g$severity == EXT_LEVEL)
            + b[["ben_major"]])   # Major benefit, US keeps pace
  v_opt <- v_opt + b[["comp_pace"]]
  p_i <- 10^((b[["no_choice"]] - v_opt) / logp_i)
  q <- quantile(p_i, c(0.1, 0.25, 0.5, 0.75, 0.9))
  tibble(severity = g$severity,
         `strictest decile` = fmt_1inN(q[1]), q25 = fmt_1inN(q[2]),
         median = fmt_1inN(q[3]), q75 = fmt_1inN(q[4]),
         `laxest decile` = fmt_1inN(q[5]))
}))
kable(het, caption = sprintf(
  "Population distribution of individual p* (mixing over the estimated random risk slope, coefficients at their point estimates; Major benefit, US keeps pace). %s of the fitted slope distribution is non-negative (no finite p*; risk-indifferent tail).",
  pct(share_nonneg)))
```

```{r dce-sensitivity}
resp_clean <- resp %>%
  filter(!d$flag_dominated[match(code, d$participant.code)] &
         !d$flag_unstable[match(code, d$participant.code)])
m_dce_sens <- fit_dce(resp_clean, "sensitivity")
sens_tab <- tibble(parameter = names(co),
                   primary = round(co, 3),
                   flags_dropped = round(coef(m_dce_sens)[names(co)], 3))
kable(sens_tab, caption = sprintf(
  "Registered sensitivity: dominance/stability-flagged respondents dropped (N = %d -> %d). Primary analysis retains them.",
  n_distinct(resp$code), n_distinct(resp_clean$code)))
```

```{r dce-wave}
# Wave robustness (§4.2): design wave as a control. All waves here fielded
# the same committed design; wave enters as an opt-out shift (a respondent-
# constant covariate is identified only against the no-choice constant).
long_all <- memo("dce_long_primary", dce_build_long(as.data.frame(resp), params$blocks_csv))
wv <- d$wave[match(long_all$code, d$participant.code)]
for (k in c("wave2", "wave3", "wave4"))
  long_all[[k]] <- as.integer(wv == k) * long_all$no_choice
m_wave <- memo("dce_wave", {
  set.seed(42)
  logitr::logitr(data = long_all, outcome = "choice", obsID = "obsID",
                 panelID = "respID",
                 pars = c(MAIN_PARS, "wave2", "wave3", "wave4"),
                 randPars = c(logp = "n"), numMultiStarts = 5, numCores = 1)
})
kable(tibble(parameter = names(coef(m_wave)),
             estimate = round(coef(m_wave), 3)),
      caption = "Wave-robustness re-estimate (wave x opt-out controls). The sequential design is ignorable for the likelihood (protocol Appendix B); substantive coefficients should be stable.")
```

# 5. Method 3 — relative standards (§4.3)

```{r m3}
m3_ids <- c("m3_std_nuclear", "m3_std_aviation", "m3_std_dams")
slab <- lab("m3_std_nuclear")
m3 <- bind_rows(lapply(m3_ids, function(id) {
  x <- num(d[[id]])
  tibble(comparator = id, response = slab,
         share = as.numeric(table(factor(x, seq_along(slab))) / sum(!is.na(x))))
}))
ggplot(m3, aes(comparator, share, fill = factor(response, rev(slab)))) +
  geom_col(position = "fill") +
  scale_fill_manual(values = c("#BFBFBF", sara_seq(5))) +  # grey = cannot compare
  scale_x_discrete(labels = c(m3_std_nuclear = "Nuclear power",
                              m3_std_aviation = "Commercial aviation",
                              m3_std_dams = "Large dams")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = NULL, fill = NULL,
       title = "How strictly should advanced AI be regulated, vs technologies we already regulate?") +
  theme_sara()
kable(m3 %>% mutate(share = pct(share)) %>%
        pivot_wider(names_from = comparator, values_from = share),
      caption = "Per-comparator distributions. 'Cannot compare these technologies' is reported as frame rejection, not missing. Registered negative commitment: NO 'N-times safer' back-out headline is computed from these items (§8).")
ev <- num(d$m3_sanity_everest)
kable(tibble(response = slab, n = as.integer(table(factor(ev, seq_along(slab))))),
      caption = sprintf("Sanity anchor (Everest, ~1 in 100 lifetime fatality stated in the stem): %s answered 'Less strict'/'Much less strict' (flagged, reported descriptively, not excluded).",
                        pct(mean(d$flag_everest, na.rm = TRUE))))
```

# 6. Method 4 — named-source tolerability (§4.4)

```{r m4}
m4_ids <- c("m2_experts_lecun", "m2_experts_fri_super",
            "m2_experts_fri_domain", "m2_experts_amodei")
m4 <- bind_rows(lapply(m4_ids, function(id) {
  x <- num(d[[id]])
  ci <- prop_ci(sum(x == 2, na.rm = TRUE), sum(!is.na(x)))
  tibble(source = id, n = sum(!is.na(x)),
         pct_intolerable = pct(ci[1]),
         ci95 = sprintf("[%s, %s]", pct(ci[2]), pct(ci[3])))
}))
kable(m4, caption = "% judging the named source's stated risk level Intolerable, BY SOURCE ONLY. Registered negative commitment: no pooling into a single 'experts say' figure — the events and horizons differ (§8).")
m4_fig <- bind_rows(lapply(m4_ids, function(id) {
  x <- num(d[[id]])
  ci <- prop_ci(sum(x == 2, na.rm = TRUE), sum(!is.na(x)))
  tibble(source = id, p = ci[1], lo = ci[2], hi = ci[3])
})) %>% mutate(source = factor(source, rev(m4_ids)))
ggplot(m4_fig, aes(p, source)) +
  geom_point(size = 2, colour = SARA_RED) +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.2, colour = SARA_RED) +
  scale_y_discrete(labels = c(
    m2_experts_lecun = "LeCun (asteroid base rate)",
    m2_experts_fri_super = "FRI superforecasters (1 in 250)",
    m2_experts_fri_domain = "FRI AI-domain experts (1 in 30)",
    m2_experts_amodei = "Amodei (1 in 10 to 1 in 4)")) +
  scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(title = "Share judging each named source's stated risk Intolerable",
       subtitle = "sources ordered near-zero (LeCun/asteroid) to top-of-range (Amodei); never pooled",
       x = "% Intolerable (95% CI)", y = NULL) +
  theme_sara()
pat <- d %>%
  transmute(across(all_of(m4_ids), ~ num(.))) %>%
  filter(if_all(everything(), ~ !is.na(.))) %>%
  count(across(everything())) %>% arrange(desc(n)) %>% head(8)
kable(pat, caption = "Most common within-person response patterns across the four sources (1 = Tolerable, 2 = Intolerable), descriptive.")
```

# 7. Disclosure arm (§4.5)

Registered contrast: each Method 1–4 headline in disclosed vs undisclosed
halves. **Where they diverge, the disclosed estimate is the headline** and
the swing is a primary finding.

```{r disclosure}
disc <- list()
for (id in ladder_ids) {
  x <- num(d[[id]])
  disc[[id]] <- tibble(
    headline = paste0("M1 median rung: ", id),
    disclosed = ladder_lab[ordinal_median(x[d$info])],
    undisclosed = ladder_lab[ordinal_median(x[!d$info])],
    p = wilcox.test(x ~ d$info)$p.value)
}
for (id in m3_ids) {
  x <- num(d[[id]])
  strict <- x %in% 1:2
  disc[[id]] <- tibble(
    headline = paste0("M3 % (much) stricter: ", id),
    disclosed = pct(mean(strict[d$info & !is.na(x)])),
    undisclosed = pct(mean(strict[!d$info & !is.na(x)])),
    p = prop.test(c(sum(strict & d$info & !is.na(x)), sum(strict & !d$info & !is.na(x))),
                  c(sum(d$info & !is.na(x)), sum(!d$info & !is.na(x))))$p.value)
}
for (id in m4_ids) {
  x <- num(d[[id]])
  intol <- x == 2
  disc[[id]] <- tibble(
    headline = paste0("M4 % intolerable: ", id),
    disclosed = pct(mean(intol[d$info & !is.na(x)])),
    undisclosed = pct(mean(intol[!d$info & !is.na(x)])),
    p = prop.test(c(sum(intol & d$info, na.rm = TRUE), sum(intol & !d$info, na.rm = TRUE)),
                  c(sum(d$info & !is.na(x)), sum(!d$info & !is.na(x))))$p.value)
}
x <- ifelse(num(d$m5_wtp) %in% dk, NA, num(d$m5_wtp))
disc$wtp <- tibble(headline = "M1 median WTP band",
                   disclosed = wlab[ordinal_median(x[d$info])],
                   undisclosed = wlab[ordinal_median(x[!d$info])],
                   p = wilcox.test(x ~ d$info)$p.value)
disc_tab <- bind_rows(disc) %>% mutate(p = signif(p, 2), diverges = p < 0.05)
kable(disc_tab, caption = "Disclosed vs undisclosed halves for every Method 1-4 headline. Rows with diverges = TRUE take the DISCLOSED value as the published headline (registered rule); the swing itself is a primary finding (RQ5).")
```

# 8. Convergence go/no-go (§4.6, registered decision rule)

A single quantitative public number is published **only if** Method 1
(stated) and Method 2 (DCE) agree within one order of magnitude in the
overlap of severity tiers, on disclosed-arm data.

```{r convergence}
resp_disc <- resp %>% filter(d$info[match(code, d$participant.code)])
m_dce_disc <- fit_dce(resp_disc, "disclosed")
# Comparison scenario (documented operationalisation): Major benefit,
# 'The US keeps pace' — the middle level of each non-risk attribute; the
# other benefit levels are shown below as gate robustness.
ps_disc_all <- pstar_draws(m_dce_disc, params$n_pstar_draws) %>%
  filter(competition == "The US keeps pace")
sev_map <- setNames(names(SEV_TO_LOGN), ladder_ids)
conv_at <- function(ben) bind_rows(lapply(ladder_ids, function(id) {
  x <- num(d[[id]])[d$info]
  med <- ordinal_median(x)
  m1_p <- if (!is.na(med) && med <= 8) 10^(-med) else NA   # Never has no numeric
  dce_p <- ps_disc_all$median[ps_disc_all$severity == sev_map[[id]] &
                              ps_disc_all$benefit == ben]
  tibble(severity = sev_map[[id]],
         m1_median = ladder_lab[med], m1_p = m1_p, dce_pstar = dce_p,
         gap_oom = ifelse(is.na(m1_p) | dce_p >= 1, NA,
                          abs(log10(m1_p) - log10(dce_p))))
}))
conv <- conv_at("Major")
kable(conv %>% mutate(m1_p = fmt_1inN(m1_p), dce_pstar = fmt_1inN(dce_pstar),
                      gap_oom = round(gap_oom, 2)),
      caption = "Method 1 (disclosed, stated) vs Method 2 (disclosed, DCE p* at Major benefit / US keeps pace). gap_oom = absolute order-of-magnitude gap; NA where a tier does not bind (p* >= 1) or the stated median is 'Never allowed'.")
binding <- conv$gap_oom[!is.na(conv$gap_oom)]
gate_pass <- length(binding) > 0 && all(binding <= 1)

gate_rob <- bind_rows(lapply(c("Modest", "Major", "Transformative"), function(b) {
  g <- conv_at(b)$gap_oom
  g <- g[!is.na(g)]
  tibble(benefit_scenario = b, binding_tiers = length(g),
         max_gap_oom = ifelse(length(g), round(max(g), 2), NA),
         verdict = ifelse(length(g) > 0 && all(g <= 1), "pass", "fail"))
}))
kable(gate_rob, caption = "Gate robustness: the same rule evaluated at each benefit scenario (the registered decision uses the Major / 'US keeps pace' scenario, per PREREGISTRATION.md §4.6).")
if (length(unique(gate_rob$verdict)) > 1) cat(
  "NOTE: the verdict differs across benefit scenarios - the registered (Major-scenario) verdict stands for the §4.6 decision, but this fragility is reported alongside any published number.\n")
```

**Gate `r if (gate_pass) "PASSES" else "FAILS"`** on the binding overlap tiers
(`r sum(!is.na(conv$gap_oom))` tiers compared).
`r if (!gate_pass) "Registered consequence: no single quantitative public number is published; the report gives a bound and qualitative findings only, and no expert-vs-public multiplier is computed (§8)."`

```{r multiplier, eval=gate_pass}
# Computed ONLY on a pass; always labelled researcher interpretation.
fri_annual <- 1 - (1 - 0.0038)^(1 / 74)   # FRI superforecaster 0.38% by 2100, annualised
pub <- conv$dce_pstar[conv$severity == EXT_LEVEL]
if (length(pub) == 1 && is.finite(pub) && pub < 1) cat(sprintf(
  "RESEARCHER INTERPRETATION (not a public-opinion finding): at the stated scenario (Major benefit, US keeps pace), the public's extinction-tier p* (%s/yr) sits about %.0fx %s the FRI superforecaster-implied annual extinction estimate (%.1e/yr).\n",
  fmt_1inN(pub), max(pub, fri_annual) / min(pub, fri_annual),
  ifelse(pub > fri_annual, "above", "below"), fri_annual))
```

# 9. Superintelligence module (§4.7)

DV construction (see the prep chunk): pro-ban top-2-box on `muskan_support`
("Don't know" missing), and the continuous score = pro-ban support minus
reverse-keyed anti-ban agreement, so acquiescence pushes the two components
in opposite directions and cancels.

### Aim 2 — considered support for the conditional ban

```{r aim2}
aim2 <- d %>% filter(!is.na(cell)) %>% group_by(cell) %>%
  summarise(n = sum(!is.na(mk_top2)),
            top2 = sum(mk_top2, na.rm = TRUE),
            anti_top2_share = pct(mean(anti_top2, na.rm = TRUE)),
            mean_score = round(mean(mk_score, na.rm = TRUE), 2)) %>%
  rowwise() %>%
  mutate(ci = { civ <- prop_ci(top2, n)
                sprintf("%s [%s, %s]", pct(civ[1]), pct(civ[2]), pct(civ[3])) }) %>%
  ungroup()
kable(aim2 %>% dplyr::select(cell, n, `pro-ban top-2-box [95% CI]` = ci,
                             anti_top2_share, mean_score),
      caption = "Ban support by cell. Cell 5 (substantive x substantive, both sides) is the registered Aim-2 headline: considered support after balanced exposure — bracketed by control (cell 9), the other contested cells, and the one-sided cells. Both DVs reported; 'Don't know' treated as missing.")
head5 <- aim2 %>% filter(cell == 5)

cell_desc <- d %>% filter(!is.na(cell)) %>% distinct(cell, muskan_for, muskan_against) %>%
  mutate(label = sprintf("%d: for=%s / against=%s", cell, muskan_for, muskan_against))
aim2_fig <- aim2 %>% left_join(cell_desc, by = "cell") %>% rowwise() %>%
  mutate(p = top2 / n, lo = prop_ci(top2, n)[2], hi = prop_ci(top2, n)[3]) %>% ungroup()
ggplot(aim2_fig, aes(p, reorder(label, cell))) +
  geom_point(aes(colour = cell == 5), size = 2, show.legend = FALSE) +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.2) +
  scale_x_continuous(labels = scales::percent) +
  scale_colour_manual(values = c(`FALSE` = "#4D4D4D", `TRUE` = SARA_RED)) +
  labs(title = "Conditional-ban support (top-2-box) by briefing cell",
       subtitle = "cell 5 (red) is the registered considered-support headline; cell 9 = no-briefing control",
       x = "pro-ban top-2-box (95% CI)", y = NULL) +
  theme_sara()
```

**Aim-2 headline (cell 5): `r head5$ci`** pro-ban top-2-box after hearing
both substantive cases.

### H2 / H3 — registered factorial with planned contrasts

```{r h2h3}
df <- d %>% filter(!is.na(mk_score))
df$For     <- factor(df$muskan_for,     levels = c("none", "elite", "substantive"))
df$Against <- factor(df$muskan_against, levels = c("none", "elite", "substantive"))
cmat <- cbind(any_vs_none = c(-2, 1, 1),        # C1
              route_elite_vs_sub = c(0, 1, -1)) # C2
contrasts(df$For) <- cmat; contrasts(df$Against) <- cmat
m_sup <- lm(mk_score ~ For * Against, data = df)   # registered model (Rev 3 plan §11)
kable(round(summary(m_sup)$coefficients, 3),
      caption = "Registered factorial: support ~ For * Against with planned contrasts (C1 any-vs-none, C2 elite-vs-substantive). H3 = the ForC1 (+) and AgainstC1 (-) main effects.")

one <- df %>% filter(cell %in% c(3, 6, 7, 8)) %>%
  mutate(direction = ifelse(cell %in% c(3, 6), "For", "Against"),
         route = ifelse(cell %in% c(3, 7), "elite", "substantive"))
m_h2 <- lm(mk_score ~ direction * route, data = one)
kable(round(summary(m_h2)$coefficients, 3),
      caption = "H2 readout (primary Aim-1 test): direction x route interaction on the one-sided cells — is the pro-vs-anti swing larger under substantive than elite messaging?")
h2_sd <- coef(m_h2)[4] / sd(one$mk_score, na.rm = TRUE)
cat(sprintf("H2 interaction in SD units: %.2f (the registered power simulation gives 0.83 power at |delta| = 0.15 SD and 0.97 at 0.20 SD with 360/cell; observed cells: %s).\n",
            h2_sd, paste(table(one$cell), collapse = "/")))
ggplot(one %>% group_by(direction, route) %>%
         summarise(m = mean(mk_score, na.rm = TRUE),
                   se = sd(mk_score, na.rm = TRUE) / sqrt(sum(!is.na(mk_score))),
                   .groups = "drop"),
       aes(route, m, group = direction, colour = direction)) +
  scale_colour_manual(values = c(For = SARA_RED, Against = "#2B2B2B")) +
  geom_point(size = 2) + geom_line() +
  geom_errorbar(aes(ymin = m - 1.96 * se, ymax = m + 1.96 * se), width = 0.08) +
  labs(title = "H2: direction x route on ban support (one-sided cells)",
       subtitle = "non-parallel lines = the pro-vs-anti swing depends on messaging route",
       x = "messaging route", y = "mean support score (95% CI)") +
  theme_sara()
```

### Quasi-manipulation-check (registered pattern)

```{r quasi-check}
one$central <- num(d$muskan_central_route[match(one$participant.code, d$participant.code)])
one$peripheral <- num(d$muskan_peripheral_route[match(one$participant.code, d$participant.code)])
t_per <- t.test(peripheral ~ route, data = one)
t_cen <- t.test(central ~ route, data = one)
bal   <- t.test(central ~ factor(cell), data = filter(one, cell %in% c(6, 8)))
qc <- tibble(
  check = c("Cue reliance higher in ELITE cells", "Elaboration higher in SUBSTANTIVE cells",
            "Elaboration balanced across cells 6 vs 8 (material balance)"),
  mean_diff = round(c(diff(rev(t_per$estimate)), diff(t_cen$estimate),
                      diff(bal$estimate)), 2),
  p = signif(c(t_per$p.value, t_cen$p.value, bal$p.value), 2))
kable(qc, caption = "Quasi-manipulation-checks (one-sided cells; self-perceived processing, stated limits per the project plan). Row 3 checks that pro and anti substantive arguments prompted equal engagement; imbalance would bias the Aim-2 contested estimate toward the stronger side.")
```

### H5′ / H6′ — mediation (EXPLORATORY)

```{r mediation}
set.seed(42)
med_boot <- function(data, mediator) {
  st <- function(dd, i) {
    dd <- dd[i, ]
    a <- coef(lm(dd[[mediator]] ~ dd$route_sub))[2]
    b <- coef(lm(dd$mk_score ~ dd[[mediator]] + dd$route_sub))[2]
    a * b
  }
  b <- boot(data, st, R = params$n_boot)
  ci <- boot.ci(b, type = "perc")$percent[4:5]
  tibble(mediator = mediator, indirect = round(b$t0, 3),
         ci95 = sprintf("[%.3f, %.3f]", ci[1], ci[2]))
}
one$route_sub <- as.integer(one$route == "substantive")
med <- bind_rows(med_boot(one[!is.na(one$central) & !is.na(one$mk_score), ], "central"),
                 med_boot(one[!is.na(one$peripheral) & !is.na(one$mk_score), ], "peripheral"))
kable(med, caption = sprintf("EXPLORATORY (registered as such): route -> mediator -> support indirect effects, one-sided cells, %d bootstrap resamples, single-item mediators (no reliability estimate possible). Not confirmatory evidence.", params$n_boot))
```

### Balance, comprehension, and module sensitivities

```{r muskan-checks}
vb <- chisq.test(table(d$cell, num(d$muskan_version, keep_optout = TRUE)))
cat(sprintf("Stimulus version balance across cells: chi-sq p = %.2f -> %s\n",
            vb$p.value,
            ifelse(vb$p.value < 0.05,
                   "IMBALANCED: version enters the models above as a random effect (registered handling)",
                   "balanced; no version term needed (registered handling)")))
comp <- num(d$muskan_si_comprehension)
correct_opt <- grep("more intelligently", lab("muskan_si_comprehension"))
cat(sprintf("Control-cell comprehension check: %s correct (flag/weight diagnostic, not an outcome).\n",
            pct(mean(comp[d$cell == 9] == correct_opt, na.rm = TRUE))))

sens_mod <- bind_rows(
  tibble(analysis = "Aim 2 (cell 5 top-2-box)",
         primary = pct(head5$top2 / head5$n),
         drop_acquiescers = pct(with(filter(d, cell == 5, !flag_acquiescer),
                                     mean(mk_top2, na.rm = TRUE))),
         drop_speeders = pct(with(filter(d, cell == 5, !flag_speed),
                                  mean(mk_top2, na.rm = TRUE)))),
  tibble(analysis = "H2 interaction coefficient",
         primary = sprintf("%.3f", coef(m_h2)[4]),
         drop_acquiescers = sprintf("%.3f", coef(update(m_h2, data = filter(one, !flag_acquiescer)))[4]),
         drop_speeders = sprintf("%.3f", coef(update(m_h2, data = filter(one, !flag_speed)))[4])))
kable(sens_mod, caption = "Registered module sensitivities: acquiescence flag (§3.5) and speed flag (§3.6) dropped (module results only; primary retains them).")
```

# 10. MRP population estimates (§6)

Demonstrated on one registered headline outcome — the share who would allow
**at most a 1-in-1,000,000 annual chance of human extinction (or never)** —
with the full registered machinery: multilevel model over age × gender ×
education × income × state, post-stratified to the ACS frame; posterior
predictive check; leave-one-state-out validation; the state-map threshold
rule. The production analysis repeats this per registered outcome.

```{r mrp}
frame_path <- if (file.exists("../poststrat_frame.csv")) "../poststrat_frame.csv" else
  file.path(DATA, "poststrat_frame_synthetic.csv")
frame <- read.csv(frame_path, check.names = FALSE)
cat(sprintf("Post-stratification frame: %s (%s)\n", frame_path,
            ifelse(grepl("synthetic", frame_path),
                   "SYNTHETIC pipeline-test frame from the simulator - real analysis uses build_poststrat.py output",
                   "ACS frame")))

md <- d %>% transmute(
  y = as.integer(num(m4c_extinction) >= 6),          # 1-in-1M or stricter, incl. Never
  age = code_to_label(demo_age, lab("demo_age")),
  gender = code_to_label(demo_gender, lab("demo_gender")),
  education = code_to_label(demo_education, lab("demo_education")),
  income = code_to_label(demo_income, lab("demo_income")),
  state = code_to_label(demo_state, lab("demo_state"))) %>%
  filter(!is.na(y),
         !grepl("Prefer not", age), !grepl("Prefer not", education),
         !grepl("Prefer not", income), !grepl("Prefer not", state))
cat(sprintf("MRP model rows: %d ('Prefer not to say' on a weighting variable: %d dropped; the inclusive gender category is retained in fitting and allocated per ACS_poststratification_manual.md)\n",
            nrow(md), sum(!is.na(num(d$m4c_extinction))) - nrow(md)))

m_mrp <- memo("mrp_glmer",
  glmer(y ~ gender + (1 | age) + (1 | education) + (1 | income) + (1 | state),
        data = md, family = binomial))
frame$gender <- frame$sex
frame$pred <- predict(m_mrp, newdata = frame, type = "response",
                      allow.new.levels = TRUE)
mrp_national <- sum(frame$pred * frame$n) / sum(frame$n)
raw <- mean(md$y)

# Approximate posterior draws for an interval: fixed effects from their
# asymptotic MVN; each random-effect level from its conditional mode and
# conditional SD. (The production analysis fits the identical model in a
# fully Bayesian framework; this is the fast pipeline approximation.)
mrp_ci <- memo("mrp_ci", {
  set.seed(42)
  ndr <- 500
  fe <- MASS::mvrnorm(ndr, fixef(m_mrp), as.matrix(vcov(m_mrp)))
  re <- ranef(m_mrp, condVar = TRUE)
  vc <- as.data.frame(VarCorr(m_mrp))
  re_draw <- function(g) {
    modes <- re[[g]][, 1]
    sds <- sqrt(as.numeric(attr(re[[g]], "postVar")))
    dr <- sapply(seq_along(modes), function(k) rnorm(ndr, modes[k], sds[k]))
    colnames(dr) <- rownames(re[[g]])
    # frame levels unseen in the sample draw from the group-level
    # distribution (standard MRP handling for empty cells)
    missing <- setdiff(unique(frame[[g]]), colnames(dr))
    if (length(missing)) {
      sd_g <- vc$sdcor[vc$grp == g][1]
      extra <- sapply(missing, function(x) rnorm(ndr, 0, sd_g))
      dr <- cbind(dr, matrix(extra, nrow = ndr,
                             dimnames = list(NULL, missing)))
    }
    dr
  }
  dr_age <- re_draw("age"); dr_edu <- re_draw("education")
  dr_inc <- re_draw("income"); dr_st <- re_draw("state")
  est <- sapply(seq_len(ndr), function(i) {
    eta <- (fe[i, "(Intercept)"]
            + ifelse(frame$gender == "Male", fe[i, "genderMale"], 0)
            + dr_age[i, frame$age] + dr_edu[i, frame$education]
            + dr_inc[i, frame$income] + dr_st[i, frame$state])
    sum(plogis(eta) * frame$n) / sum(frame$n)
  })
  quantile(est, c(0.025, 0.975))
})
kable(tibble(estimate = c("Raw sample share", "MRP population estimate"),
             value = c(pct(raw),
                       sprintf("%s [95%%: %s, %s]", pct(mrp_national),
                               pct(mrp_ci[1]), pct(mrp_ci[2])))),
      caption = "MRP estimate with an approximate 95% interval (draw-based; see code). Read next to the §7 benchmark gaps: MRP adjusts demographics only, not panel selection.")
```

```{r mrp-checks, fig.height=4}
# Posterior-predictive-style check: model-implied vs observed state shares
obs <- md %>% group_by(state) %>% summarise(obs = mean(y), n = n())
pred_state <- frame %>% group_by(state) %>%
  summarise(pred = sum(pred * n) / sum(n))
ppc <- inner_join(obs, pred_state, by = "state")
ggplot(ppc, aes(pred, obs, size = n)) +
  geom_point(alpha = .6, colour = SARA_RED) +
  geom_abline(linetype = 2, colour = SARA_GREY) + theme_sara() +
  labs(title = "Predictive check: model-implied vs observed state shares",
       x = "post-stratified model prediction", y = "raw state share")

# Leave-one-state-out validation (fast refits)
loso <- memo("mrp_loso", {
  bind_rows(lapply(unique(md$state), function(s) {
    m <- suppressMessages(glmer(
      y ~ gender + (1 | age) + (1 | education) + (1 | income) + (1 | state),
      data = filter(md, state != s), family = binomial, nAGQ = 0))
    fs <- filter(frame, state == s)
    tibble(state = s,
           pred = sum(predict(m, fs, type = "response",
                              allow.new.levels = TRUE) * fs$n) / sum(fs$n),
           obs = mean(md$y[md$state == s]), n = sum(md$state == s))
  }))
})
cat(sprintf("Leave-one-state-out: mean absolute error %.1f pp across %d states.\n",
            100 * mean(abs(loso$pred - loso$obs)), nrow(loso)))

# Registered state-map threshold
kable(head(arrange(obs, desc(n)), 5),
      caption = sprintf("Largest per-state samples. Registered rule: state maps are published only where the effective per-state sample exceeds 300. Max here = %d -> %s",
                        max(obs$n),
                        ifelse(max(obs$n) > 300, "maps allowed for qualifying states",
                               "NO state map is published for this outcome")))
```

# 11. Registered negative commitments (§8) — compliance

```{r commitments}
kable(tibble(
  commitment = c(
    "No 'N times safer' multiplier from the Method 3 back-out; any multiplier labelled researcher interpretation",
    "No pooling of Method 4 sources",
    "No point estimate of p* — distributions with intervals only",
    "No state map below the effective-sample threshold",
    "No mid-wave design changes; no within-respondent DCE adaptation",
    "No re-labelling of exploratory analyses as confirmatory"),
  status = c(
    "complied - §5 reports distributions only; §8's multiplier (if any) is gated and labelled",
    "complied - §6 reports by source only",
    "complied - §4 reports medians with 95% intervals",
    sprintf("complied - §10 threshold rule applied"),
    "complied - single committed design per wave (md5 in §0); estimation on tasks 1-8 only",
    "complied - H5'/H6' and proxy moderators labelled exploratory throughout")),
  caption = "Checklist against PREREGISTRATION.md §8.")
```

`r if (SIMULATED) '# 12. Parameter recovery (simulated data only)'`

```{r recovery, eval=SIMULATED}
tb <- unlist(truth$truth$dce)
est_p <- coef(m_dce); est_s <- coef(m_dce_sens)[names(est_p)]
# logitr's random-normal SD is identified only up to sign
est_p["sd_logp"] <- abs(est_p["sd_logp"]); est_s["sd_logp"] <- abs(est_s["sd_logp"])
rec <- tibble(parameter = names(est_p), truth = tb[names(est_p)],
              primary = round(est_p, 3),
              err_primary = round(abs(est_p - tb[names(est_p)]), 3),
              flags_dropped = round(est_s, 3),
              err_flags = round(abs(est_s - tb[names(est_p)]), 3))
kable(rec, caption = sprintf(
  "DCE: registered mixed logit vs simulation ground truth (compare the pre-freeze identification runs in PREREGISTRATION.md §5b). The DGP deliberately includes %.0f%% random-choice respondents, so the PRIMARY estimate is attenuated toward zero by design; the flags-dropped sensitivity removes most of them (a random chooser fails the dominance or stability check with probability 7/9), so recovery is judged there.",
  100 * truth$truth$behaviour$p_dce_random))
ok <- all(rec$err_flags < 0.15, na.rm = TRUE)
cat(if (ok) "RECOVERY OK: every DCE parameter within 0.15 of truth on the flags-dropped estimate; primary shows the expected contamination attenuation.\n" else
    "RECOVERY WARNING: at least one DCE parameter is off by >= 0.15 on the flags-dropped estimate - inspect before trusting the pipeline.\n")
# Superintelligence module: the truth orders the one-sided cells
# for-substantive > for-elite > control > against-elite > against-substantive
# on the latent scale; the observed mean scores must reproduce that ranking
# and a positive H2 interaction (the against side moves more under substance).
mt <- truth$truth$muskan
mk_rec <- tibble(cell = c(6, 3, 9, 7, 8),
                 arm = c("for-substantive", "for-elite", "control",
                         "against-elite", "against-substantive"),
                 truth_latent = c(mt$for_substantive, mt$for_elite, 0,
                                  mt$against_elite, mt$against_substantive)) %>%
  left_join(d %>% filter(!is.na(cell)) %>% group_by(cell) %>%
              summarise(observed_mean_score = round(mean(mk_score, na.rm = TRUE), 2)),
            by = "cell")
kable(mk_rec, caption = "Superintelligence module: truth (latent) vs observed mean support score by arm. Cells 3 and 6 differ by only 0.05 latent units in the truth — indistinguishable at ~400/cell by design — so recovery requires the detectable ordering (both For arms above control, control above both Against arms, against-elite above against-substantive), not the full ranking.")
rho <- cor(mk_rec$truth_latent, mk_rec$observed_mean_score, method = "spearman")
sc <- setNames(mk_rec$observed_mean_score, mk_rec$arm)
rank_ok <- all(sc[c("for-substantive", "for-elite")] > sc["control"],
               sc["control"] > sc[c("against-elite", "against-substantive")],
               sc["against-elite"] > sc["against-substantive"])
# matches the lm parametrisation: (For,sub - For,elite) - (Against,sub - Against,elite)
h2_truth <- (mt$for_substantive - mt$for_elite) -
            (mt$against_substantive - mt$against_elite)
cat(sprintf("Detectable cell ordering recovered: %s (Spearman rho = %.1f). Simulated H2 interaction (latent): %.2f; observed interaction coefficient: %.2f (same sign: %s).\n",
            ifelse(rank_ok, "YES", "NO - inspect"), rho,
            h2_truth, coef(m_h2)[4], sign(h2_truth) == sign(coef(m_h2)[4])))

# Disclosure arm: a true strictness shift was simulated in the disclosed
# half for the M1/M3/M4/WTP outcomes - §7 should detect it broadly.
cat(sprintf("Info-arm truth: latent strictness shift of %.2f SD was simulated. §7 detected divergence (p < .05) on %d of %d headlines.\n",
            truth$truth$info$strict_shift, sum(disc_tab$diverges), nrow(disc_tab)))
cat(sprintf("Ladder truth: %.0f%% simulated scope-flat at mean numeracy (observed overall: %s); 'never at extinction' share %.0f%% (observed 'Never allowed' at extinction: %s).\n",
            100 * plogis(truth$truth$ladder$flat_intercept),
            pct(mean(d$scope_flat, na.rm = TRUE)),
            100 * truth$truth$ladder$never_ext_share,
            pct(mean(rungs[, "m4c_extinction"] == 9, na.rm = TRUE))))
```

# `r if (SIMULATED) '13' else '12'`. Session info

```{r session}
sessionInfo()
```
